% =============================================================================
% Discussion + Conclusion + Acknowledgments + Author Contributions + Conflicts
% of Interest + Funding + AI Use Disclosure + Data Availability Statement +
% Ethics Statement. Compiles via the canonical wrapper; see BUILD.md.
% =============================================================================


\section{Discussion}
\label{sec:discussion}

\subsection{Principal findings}
\label{discussion:findings}

Aggregate AUC reporting concealed within-stratum AUCs near the no-discrimination floor, with paradox gaps to 16.5 pp in the worst-affected preoperative model. Cross-population transferability ran 8.5 pp better in one direction than the other, persisted at roughly 70\% of baseline after multi-dimensional case-mix matching, and was invisible to any single-direction validation. Pre-Platt calibration slopes spanned 0.41--1.29 across all eight cross-population runs and recovered to 0.95--1.02 only after 5-fold cross-validated Platt scaling. Each of these magnitudes is clinically material at the patient-decision threshold and would not have surfaced under conventional aggregate evaluation. The methodological commitments that surface them are individually well-established TRIPOD+AI-aligned practices: stratified within-cohort analysis across clinically meaningful risk strata; bidirectional testing with multi-dimensional matched case-mix sensitivity; and case-level paired bootstrap inference with rank-invariant metrics reported separately from probability-scale metrics. The sections that follow position each finding against the literature (\S\ref{discussion:literature}), examine the underlying mechanisms (\S\ref{discussion:mechanisms}), derive methodological implications (\S\ref{discussion:deployment}), and consider the constraints that limit our conclusions (\S\ref{discussion:limitations}).

\subsection{Comparison with existing literature}
\label{discussion:literature}

Recent perioperative-medicine systematic review~\cite{arina2024probast} reports that only 13\% of published machine-learning prediction models undergo external validation, and that single-center development dominates the literature; PROBAST assessments identify high risk of bias in most models. Within this landscape, the magnitudes we report (paradox gaps to 16.5 pp, direction asymmetry of 8.5 pp persisting at $\sim$70\% of baseline after matched-case-mix sensitivity, calibration slope ranges from 0.41 to 1.29 across cross-population runs) characterize what survives bidirectional testing and stratified evaluation in the rare case that both can be applied. To our knowledge, paradox-gap magnitudes of this scale have not previously been reported in the deployed clinical prediction model literature; Chipman and Braun~\cite{chipman2017simpson} established the analytic basis for stratum-level miscalibration producing paradox in performance metrics, and our findings extend that pattern to AUC in cross-continental perioperative deployment.

The direction-asymmetry finding extends a smaller cross-population literature: Futoma et al.~\cite{Futoma2020} and Finlayson et al.~\cite{Finlayson2021} characterized the ``myth of generalisability'' and distributional shift as core deployment risks but did not quantify direction-of-transfer asymmetry or provide a case-level paired bootstrap framework. The SHAP transferability finding (Spearman $\rho = 0.93$--$0.97$) characterizes feature-importance stability between geographically separated academic medical centers; prior SHAP-robustness work focused on within-dataset stability~\cite{Lundberg2017}, and the cross-population result suggests that the relational structure of what models learn is more robust than the absolute calibration on probability scale.

\subsection{Mechanisms}
\label{discussion:mechanisms}

The four-dimension matched-sensitivity analysis clarifies the mechanism underlying the observed direction asymmetry: ASA stratum matching amplifies the asymmetry to 128--203\% of baseline; Elixhauser comorbidity matching preserves or amplifies it (113--126\%); emergency-case proportion attenuates it to 70--86\% but does not eliminate it; and temporal period was untestable due to INSPIRE timestamp anonymization. The residual asymmetry surviving the three testable matches --- approximately 70\% of baseline --- is attributable to training-population characteristics beyond measured case-mix: healthcare-system-specific feature--outcome relationships, within-stratum severity coding that ASA and Elixhauser do not capture, and unmeasured temporal effects. At matched ASA stratum, INSPIRE deaths spread across the acuity spectrum (47.2\% / 52.8\% in ASA 1--2 / $\geq 3$) while MOVER deaths concentrate in high-acuity patients (1.1\% / 98.9\%) --- a 50.6-percentage-point split-difference consistent with cross-institutional feature--outcome relationships differing beyond what stratum matching equalizes. ASA physical status itself shows substantial inter-rater variability across institutions and clinicians~\cite{Sankar2014}, so a ``severity-3'' patient under INSPIRE coding may not be clinically equivalent to a ``severity-3'' patient under MOVER coding. Cross-institutional surveys of Asian-ICU end-of-life decision-making have documented systematic differences between Korean and Western practice patterns~\cite{Phua2015,Park2018}, consistent with the healthcare-system-specific mechanism interpretation above. Supplementary \S\ref{supp:s7-mechanism-evidence} reports stratified mortality distributions and the supporting cross-institutional literature.

That ASA stratum, comorbidity burden, and emergency-case proportion all leave most of the asymmetry intact, and that ASA matching in fact amplifies it, is the strongest evidence we can offer that the finding is not an artifact of case-mix mismatch. The temporal dimension we could not test directly; INSPIRE's privacy-preserving relative timestamps foreclose calendar-aligned matching. We name this gap because it sets a real boundary on confidence: the residual asymmetry could in principle reflect temporal-period effects we have no way to estimate. The finding stands; future work benefits from cohorts where calendar-date alignment is available.

The SHAP-plus-calibration finding provides a mechanistic lens on what transfers. Feature importance rankings are preserved almost perfectly ($\rho \geq 0.93$): a Seoul-trained model evaluated on Irvine data ranks feature importance (ASA, BMI, HR dynamics) in the same order as on its training cohort, yet absolute risk probabilities are systematically biased and require recalibration. Calibration fails uniformly before Platt scaling (O:E 0.02 to 2.92; slopes 0.41 to 1.29). This dissociation --- discrimination is preserved as a ranking, but calibration is distorted on the probability scale --- aligns with the well-documented pattern that discrimination transfers more readily than calibration~\cite{VanCalster2019}, supporting the interpretation that cross-population models preserve relational structure (which features matter, in what order) but not absolute mappings (what probability a given feature vector implies).

Returning to the asymmetry finding, this case study cannot directly answer how the pattern might run with training drawn from an LMIC tertiary perioperative cohort. Healthcare-system-specific feature--outcome relationships, named in the residual-mechanism analysis above as one component of the asymmetry that case-mix matching cannot equalize, are especially relevant in this context. LMIC tertiary perioperative care chains differ from HIC counterparts in ways case-mix variables do not capture: referral patterns, selection-into-surgery under limited non-operative alternatives, and post-operative monitoring infrastructure all shape feature--outcome relationships in systematically different ways. Whether models trained on LMIC perioperative cohorts would transfer to HIC deployment more or less successfully than the conventional HIC$\to$LMIC direction is not something present data can settle; paired LMIC--HIC perioperative datasets at INSPIRE or MOVER scale do not yet exist. The asymmetry mechanism makes the question relevant for deployment contexts where consequences are largest.

\subsection{Clinical implications}
\label{discussion:deployment}

These findings motivate three concrete changes to current practice. First, stratified external validation within clinically meaningful risk groups (e.g., ASA 1--2 vs $\geq 3$; elective vs emergency) should become standard for TRIPOD+AI-compliant reporting; aggregate AUC should accompany, not replace, within-stratum AUCs and paradox gaps (Supplementary Table~\ref{tab:paradox_summary}). Second, cross-population claims should include bidirectional transfer testing with multi-dimensional matched-case-mix sensitivity; $A \to B$ without reciprocal $B \to A$ cannot detect direction asymmetry, and matching is required to distinguish genuine failure from case-mix mismatch. Third, cross-population deployment requires mandatory probability recalibration (Platt or isotonic) on a representative sample of the deployment population~\cite{eylerdang2021right}. Recent external-validation sample-size guidance suggests that several thousand cases are typically needed to estimate calibration slope and O:E precisely, depending on the anticipated event proportion and degree of miscalibration~\cite{Riley2021Ext}. Discrimination-focused reporting alone is insufficient; published AUCs should accompany calibration assessments and recalibration protocols.

These methodological commitments are particularly consequential for HIC$\to$LMIC deployment of clinical prediction models, where healthcare-system-specific feature--outcome relationships (\S\ref{discussion:mechanisms}) and the absence of paired LMIC--HIC perioperative datasets at scale together constrain the evidence base reviewers and deployers can draw on; the cautionary patterns we report (direction asymmetry, calibration collapse, paradox gaps) are precisely the failure modes such deployments face.

\subsection{Limitations}
\label{discussion:limitations}

Five principal limitations apply. First, only two institutions (Seoul National and UC Irvine) were included, characterizing cross-institutional rather than cross-national transfer. Second, temporal-period matching was infeasible due to INSPIRE's privacy-preserving relative-time offsets; this constraint and its implications for the residual asymmetry are addressed in \S\ref{discussion:mechanisms}. Third, neither race/ethnicity stratification (omitted from the public MOVER release; no Korean comparator) nor within-stratum severity matching beyond Elixhauser comorbidity could be uniformly operationalized; race/ethnicity is therefore not analyzed. This prevents detection of subgroup discrimination failures along race or ethnicity --- a well-documented fairness failure mode in cross-population deployment~\cite{Chen2020,Pfohl2021}. Fourth, procedure-type matching was not attempted because INSPIRE's KCD-10 and MOVER's CPT/EPIC code taxonomies do not align without a purpose-built crosswalk. Fifth, the 60-minute post-induction window was chosen by clinical convention; alternative lengths were not systematically examined. These constraints bound the empirical reach of the present case study but do not affect its standing as a cautionary reference for cross-population evaluation practice.

\section{Conclusion}
\label{sec:conclusion}

This case study reports cross-continental perioperative mortality prediction failure modes at clinically material magnitudes. We position these findings as a cautionary reference for cross-population deployment of clinical prediction models, in the lineage of post-deployment evaluation work in clinical machine learning~\cite{wong2021epic}. Trained model objects, analysis code, and aggregate verification outputs are archived for reproducibility (\S\ref{methods:reproducibility}). Paired LMIC--HIC perioperative cohorts with calendar-aligned timestamps are a natural next step.

\section*{Acknowledgments}
The author acknowledges the Seoul National University Hospital INSPIRE team for stewardship of the INSPIRE dataset~\cite{Lim2024} and the UC Irvine perioperative team for stewardship of the MOVER dataset~\cite{Samad2023}.

\section*{Author Contributions}
Sole author. The author was responsible for all aspects of the work, including study conception and design, data acquisition under credentialed-access agreements, software implementation, formal analysis, interpretation, manuscript drafting, and final approval. The CRediT taxonomy roles --- conceptualization, methodology, software, validation, formal analysis, investigation, resources, data curation, writing (original draft), writing (review and editing), visualization, supervision, project administration, and funding acquisition --- all apply to the sole author.

\section*{Conflicts of Interest}
The author declares no competing interests.

\section*{Funding}
This work received no external funding.

\section*{AI Use Disclosure}
Anthropic's Claude (Claude Sonnet 4 and Claude Opus 4 series, accessed via Claude Code and the Claude.ai web interface, 2025--2026) was used during preparation of this manuscript for two purposes: (i) generation and refinement of analysis and figure-generation code in Python, which was reviewed and validated by the author before use; and (ii) editorial assistance with prose revision, copyediting, and structural organization of author-written drafts. All scientific claims, statistical analyses, study design choices, and interpretations are the author's; every quantitative result reported was independently validated against source data and primary literature by the author. No AI system is listed as an author, consistent with COPE and ICMJE guidance that AI tools cannot meet authorship criteria.

\section*{Data Availability Statement}
INSPIRE is publicly available via PhysioNet under credentialed access~\cite{INSPIRE_dataset}; MOVER is publicly available via the UC Irvine Machine Learning Repository under credentialed access~\cite{MOVER_dataset}. Both datasets were accessed under their respective data-use agreements; no patient-level data are redistributed in this paper. Analysis code, figure-generation scripts, trained model objects, and aggregate analysis outputs are archived at Zenodo (DOI: 10.5281/zenodo.19941764; see Methods \S\ref{methods:reproducibility}). Per-case prediction outputs are not redistributed (data-use agreement constraint); credentialed researchers can reproduce all analyses by applying the released models to source data after re-credentialing on PhysioNet and the UCI Machine Learning Repository.

\section*{Ethics Statement}
This study used de-identified data released for research use under the data-use agreements of INSPIRE (PhysioNet) and MOVER (UC Irvine Machine Learning Repository). No additional Institutional Review Board review was sought, as the work involves only secondary analysis of publicly released, de-identified datasets accessed under credentialed-use agreements. The author followed the data-use terms of both repositories.

% =============================================================================
% END OF DISCUSSION
% =============================================================================
